---
title: "TSCR-MCC"
output: html_notebook
---

```{r}
suppressMessages({
  library(Seurat)
  library(ggplot2)
  library(future)
  library(RColorBrewer)
  library(SingleCellExperiment)
  library(BayesSpace)
  library(dplyr)
})
```

```{r}
options(future.globals.maxSize = 50*1024*1024^2)
set.seed(100)
`%!in%` <- Negate(`%in%`)
source("utils/plot_func.R")

mp <- readxl::read_excel("~/db/41586_2023_6130_MOESM6_ESM.xlsx", sheet = 1)
mp <- lapply(mp, c)
names(mp) <- gsub("[ ]+"," ",names(mp))
```

# Objective

Identify differences in tumor purity, tumor cell heterogeneity and tumor cell characteristics (metabolic characteristics, stemness, EMT, neuroblastic and epithelial differentiation), but also signs of cells stress or invasive /migratory potential in correlation to TME composition (focus on immune [T and B cells, monocytes, macrophages and DC], mesenchymal cells [fibroblasts, endothelial cells), but also epithelial cells.

For each scRNAseq data, look for possible interactions and mutual modulations.

Use spatial data to look for tumor subpopulations in special areas (e.g. hypoxic metabolism close to the epidermis, invasive characteristics close to the invasive front or blood/lymphatic vessels, cell stress close to immune cells).

Look for signs of tumor evolution, modulated by TME. Combine scRNAseq pseudotime/velocyto with spatial data.

Associate the results with clinical course.

Read scRNA data from disk

```{r}
files <- list.files("~/data/mcc/seurat.outs/", full.names = T)
mcc.sc <- lapply(files, readRDS)
unlist(lapply(mcc.sc, function(so){unique(as.character(so$orig.ident))}))
#  [1] "D19-101"           "DD23-1035_Gex"     "E21-1-342"         "E21-1-343"         "E21-1-344"         "E23-1-345"         "E23-1-367-MCC"    
#  [8] "E23-1-384"         "E24-1-393_Gex"     "MCC01"             "MCC02"             "MCC04"             "MCC05"             "MCC07"            
# [15] "MCC08"             "MCC09"             "MCC10"             "MCC12"             "MCC13"             "MCC14"             "MCC15"            
# [22] "MCC16"             "MCC18"             "U23-1-378-MCC-NEW"
mcc.sc <- mcc.sc[-22] # Remove MCC16 (SCLC)
```

Patient/sample ID mapping

```{r}


tscrid <- c("MCC01"="BO19-1-100",
            "MCC02"="E19-1-100",
            "MCC03"="E19-1-101",
            "MCC04"="E19-1-102",
            "MCC05"="BO19-1-101",
            "MCC06"="E19-1-103",
            "MCC07"="BO19-1-102",
            "MCC08"="BO19-1-103",
            "MCC09"="BO19-1-104",
            "MCC10"="BO19-1-106",
            "MCC11"="BO19-1-107",
            "MCC12"="E19-1-104",
            "MCC13"="E19-1-108",
            "MCC14"="BO19-1-108",
            "MCC15"="BO19-1-109",
            "MCC16"="BO19-1-110",
            "MCC18"="BO19-1-111",
            "In progress"="BO20-1-112",
            "D19-101"="D19-1-101",
            "DD23-1035_Gex"="D23-1-351",
            "In progress"="DU24-1-400",
            "In progress"="E21-1-106",
            "In progress"="E21-1-109",
            "E21-1-342"="E21-1-342",
            "E21-1-343"="E21-1-343",
            "E21-1-344"="E21-1-344",
            "E23-1-345"="E23-1-345",
            "E23-1-367-MCC"="E23-1-367",
            "E23-1-384"="E23-1-384",
            "E24-1-393_Gex"="E24-1-393",
            "In progress"="U23-1-378",
            "U23-1-375_37989"="U23-1-375"
            )

patient.code <- c("E19-1-101"="811567",
                  "E19-1-102"="814642",
                  "E19-1-103"="811567",
                  "E19-1-104"="811567",
                  "E19-1-108"="811567",
                  "E24501/19 1"="811567",
                  "D2143/20 1J"="811567",
                  "BO19-1-101"="-", # E
                  "BO19-1-102"="-", # E
                  "BO19-1-103"="19281449",
                  "BO19-1-104"="19307283",
                  "BO19-1-106"="-",
                  "BO19-1-108"="19464388",
                  "BO19-1-109"="19476816",
                  "BO19-1-110"="-",
                  "BO19-1-111"="-",
                  "BO20-1-112"="20389201",
                  "BO20-1-113"="-",
                  "E21-1-342"="1981782",
                  "E21-1-343"="1981782",
                  "E21-1-344"="1981782",
                  "E21-1-106"="886503",
                  "E21-1-109"="886503",
                  "E23-1-345"="2051025",
                  "E23-1-367"="2051025",
                  "E23-1-384"="2131941",
                  "E24-1-393"="1981782",
                  "D23-1-351"="344",
                  "U23-1-375"="-",
                  "U23-1-378"="32365402",
                  "DU24-1-400"="-",
                  "22TSCR-0575"="283507",
                  "22TSCR-0694"="798314",
                  "22TSCR-0689"="772389",
                  "22TSCR-0695"="798920",
                  "22TSCR-0614"="MCC TRIM",
                  "22TSCR-0572"="577169",
                  "22TSCR-0538"="444450"
                  )

date.incoming <- c("E19-1-102"="17.01.2019",
                   "E19-1-101"="17.01.2019",
                   "E19-1-103"="11.02.2019",
                   "E19-1-104"="25.03.2019",
                   "E19-1-105"="25.03.2019",
                   "E19-1-108"="02.07.2019",
                   "E21-1-106"="25.01.2021",
                   "E21-1-109"="11.03.2021",
                   "E21-1-342"="14.12.2021",
                   "E21-1-343"="22.12.2021",
                   "E21-1-344"="17.01.2022",
                   "E23-1-345"="13.01.2023",
                   "E23-1-367"="17.03.2023",
                   "E23-1-384"="21.11.2023",
                   "E24-1-393"="24.04.2024",
                   "D23-1-351"="07.02.2023",
                   "DU24-1-400"="21.09.2023",
                   "BO19-1-101"="18.01.2019",
                   "BO19-1-102"="11.03.2019",
                   "BO19-1-103"="15.03.2019",
                   "BO19-1-104"="24.04.2019",
                   "BO19-1-106"="17.05.2019",
                   "BO19-1-108"="17.07.2019",
                   "BO19-1-109"="06.08.2019",
                   "BO19-1-110"="02.09.2019",
                   "BO19-1-111"="27.11.2019",
                   "BO20-1-112"="05.03.2020",
                   "BO20-1-113"="11.12.2020",
                   "U23-1-375"="10.05.2023",
                   "U23-1-378"="21.09.2023"
                   )

# PT or Met/VP or VN/Therapy/Response
sample.status <- c("E19-1-101"="PT/VP/BT/-",
                  "E19-1-102"="Met/VP/Naive/-",
                  "E19-1-103"="Met/VP/Avelumab SOT/NR",
                  "E19-1-104"="Met/VP/Avelumab 6w/MR",
                  "E19-1-108"="Met/VP/Nivolumab SOT/R",
                  "E24501/19 1"="Met/VP/Nivolumab 8w/R",
                  "D2143/20 1J"="Met/VP/Nivolumab 10m/R",
                  "BO19-1-101"="-/VP/-/-", # E
                  "BO19-1-102"="-/VP/-/-", # E
                  "BO19-1-103"="-/VP/BT/R",
                  "BO19-1-104"="PT/VN/-/-",
                  "BO19-1-106"="-/VP/-/-",
                  "BO19-1-108"="-/VP/-/-",
                  "BO19-1-109"="-/VP/-/-",
                  "BO19-1-110"="-/VN/-/-",
                  "BO19-1-111"="-/VP/-/-",
                  "BO20-1-112"="-/VP/-/-",
                  "BO20-1-113"="Met/VP/-/-",
                  "E21-1-342"="Met/VP/BT/NR",
                  "E21-1-343"="Met/VP/1w/NR",
                  "E21-1-344"="Met/VP/5w/R",
                  "E21-1-106"="-/-/BT/-",
                  "E21-1-109"="-/-/8w/-",
                  "E23-1-345"="Met/-/BT/-",
                  "E23-1-367"="Met/-/8w/-",
                  "E23-1-384"="-/-/BT/R",
                  "E24-1-393"="Met/VP/Avelumab BT/-",
                  "D23-1-351"="LN Met/-/Naive/-",
                  "U23-1-375"="Met/-/-/-",
                  "U23-1-378"="-/-/BT/NR",
                  "DU24-1-400"="-/-/Naive/-",
                  "22TSCR-0575"="PT/VN/-/-",
                  "22TSCR-0694"="PT/VP/-/-",
                  "22TSCR-0689"="Unclear/VP/-/-",
                  "22TSCR-0695"="PT/VN/-/-",
                  "22TSCR-0614"="PT/VN/-/-",
                  "22TSCR-0572"="PT/VP/-/-",
                  "22TSCR-0538"="PT/VN/-/-"
                  )

# tcr <- c("BO19-1-100"=113, # T cell numbers
#          "BO19-1-101"=-1,
#          "BO19-1-102"=37,
#          "BO19-1-103"=88,
#          "BO19-1-104"=389,
#          "BO19-1-106"=350,
#          "BO19-1-111"=3654,
#          "BO19-1-100"=113,
#          "E21-1-342"=75,
#          "E21-1-343"=35,
#          "E21-1-344"=372,
#          "U23-1-378"=42,
#          "E23-1-384"=49,
#          "E24-1-393"=1099,
#          "E19-1-102"=-1,
#          "E19-1-104"=-1,
#          "E19-1-108"=-1
#          )

tcr <- c("BO19-1-100"=113, # T cell numbers
         "BO19-1-101"=54,
         "BO19-1-102"=37,
         "BO19-1-103"=88,
         "BO19-1-104"=389,
         "BO19-1-106"=350,
         "BO19-1-109"=46,
         "BO19-1-110"=21,
         "BO19-1-111"=3654,
         "BO20-1-112"=46,
         "BO20-1-113"=754,
         "E21-1-342"=79,
         "E21-1-343"=46,
         "E21-1-344"=382,
         "U23-1-375"=0,
         "U23-1-378"=1314,
         "E19-1-101"=0,
         "E19-1-102"=389,
         "E19-1-103"=10,
         "E19-1-104"=25,
         "E19-1-108"=202,
         "E21-1-106"=633,
         "E21-1-109"=1118,
         "E23-1-345"=166,
         "E23-1-367"=78,
         "E23-1-384"=49,
         "E24-1-393"=1099,
         "D23-1-351"=0
         
         )

# bcr <- c("BO19-1-103"=45, # B cell numbers
#          "BO19-1-106"=70,
#          "BO19-1-108"=20,
#          "U23-1-378"=38,
#          "E23-1-384"=49,
#          "E24-1-393"=13691
#          )

bcr <- c("E19-1-101"=1,
         "E19-1-102"=354,
         "E19-1-103"=11,
         "E19-1-104"=3,
         "E19-1-108"=32,
         "E21-1-106"=61,
         "E21-1-109"=14,
         "E21-1-342"=1808,
         "E21-1-343"=240,
         "E21-1-344"=1206,
         "E23-1-345"=104,
         "E23-1-367"=47,
         "E23-1-384"=49,
         "E24-1-393"=13691,
         "D23-1-351"=5,
         "BO19-1-101"=1,
         "BO19-1-102"=3,
         "BO19-1-103"=51,
         "BO19-1-106"=82,
         "BO19-1-108"=20,
         "BO19-1-109"=33,
         "BO19-1-110"=2,
         "BO19-1-111"=2545,
         "BO20-1-112"=118,
         "BO20-1-113"=16,
         "U23-1-375"=0,
         "U23-1-378"=73
         )


fixed <- c("E19-1-104"="T019-0021",
           "E19-1-103"="T19-0012",
           "E19-1-108"="D3394-19-1B",
           "D2143/20 1J"="D2143-2B1J",
           "E24501/19 1"="E24501-19-1")

spatial.cryo <- c("E19-1-108"="P25257_101",
                  "BO19-1-109"="P25257_201",
                  "BO19-1-101"="P25257_301",
                  "BO20-1-112"="P25257_401")

spatial.ffpe <- c("E19-1-101"="D305-19",
                  "E19-1-103"="T19-0012",
                  "E19-1-108"="D3390-19",
                  "22TSCR-0575"="7632_a1",
                  "Me31"="Me31",
                  "22TSCR-0694"="60352_22-2b",
                  "22TSCR-0689"="30540_22",
                  "22TSCR-0695"="2297_23_lie",
                  "22TSCR-0614"="21-44883D",
                  "22TSCR-0572"="E2021-24662-1c",
                  "22TSCR-0538"="1489II-21")
```

```{r}
# samples.meta <- data.frame("sampleid"=unlist(lapply(mcc.sc, function(so){unique(as.character(so$orig.ident))})), "scRNA"=TRUE, row.names = unlist(lapply(mcc.sc, function(so){unique(as.character(so$orig.ident))})))
samples.meta <- as_tibble(as.data.frame(patient.code), rownames = "sample.histo")
samples.meta <- samples.meta %>% mutate(date.income=plyr::mapvalues(sample.histo, from=names(date.incoming), to=date.incoming)) %>% mutate(date.income=case_when(date.income==sample.histo ~ NA, TRUE ~ date.income))
samples.meta <- samples.meta %>% mutate(clinical = plyr::mapvalues(sample.histo, from=names(sample.status), to=sample.status))
samples.meta.clinical <- as_tibble(t(as.data.frame(strsplit(samples.meta$clinical,"/"))))
colnames(samples.meta.clinical) <- c("loc","virus","treatment","response")
samples.meta <- cbind(samples.meta, samples.meta.clinical)
samples.meta$clinical <- NULL


rownames(samples.meta) <- plyr::mapvalues(rownames(samples.meta), from=names(tscrid), to=tscrid)
rownames(samples.meta) <- gsub("_Gex","",rownames(samples.meta))
rownames(samples.meta) <- gsub("-MCC[A-Z\\-]*","",rownames(samples.meta))
# samples.meta <- samples.meta %>% mutate(patient=plyr::mapvalues(samples.meta$sample.histo, from=names(patient.code), to=patient.code)) %>% mutate(patient=case_when(patient==rownames(samples.meta) ~ NA, TRUE ~ patient))
samples.meta <- samples.meta %>% mutate(scrna=sample.histo) %>% mutate(scrna=case_when(scrna %in% tscrid ~ scrna, TRUE~NA)) %>% mutate(scrna=plyr::mapvalues(scrna, from=tscrid, to=names(tscrid)))
samples.meta <- samples.meta %>% mutate(tcr=plyr::mapvalues(sample.histo, from=names(tcr), to=tcr), bcr=plyr::mapvalues(sample.histo, from=names(bcr), to=bcr)) %>% mutate(tcr=case_when(tcr==sample.histo ~ NA, TRUE ~ tcr), bcr=case_when(bcr==sample.histo ~ NA, TRUE ~ bcr))
samples.meta <- samples.meta %>% mutate(snRNA=plyr::mapvalues(sample.histo, from=names(fixed), to=fixed)) %>% mutate(snRNA=case_when(snRNA==sample.histo ~ NA, TRUE ~ snRNA))

samples.meta <- samples.meta %>% mutate(spatial.cryo=plyr::mapvalues(sample.histo, from=names(spatial.cryo), to=spatial.cryo)) %>% mutate(spatial.cryo=case_when(spatial.cryo==sample.histo ~ NA, TRUE ~ spatial.cryo))
samples.meta <- samples.meta %>% mutate(spatial.ffpe=plyr::mapvalues(sample.histo, from=names(spatial.ffpe), to=spatial.ffpe)) %>% mutate(spatial.ffpe=case_when(spatial.ffpe==sample.histo ~ NA, TRUE ~ spatial.ffpe))
```

```{r}
samples.meta %>% arrange(patient.code)
write.csv(samples.meta %>% arrange(patient.code), "~/data/mcc/casereport_811547/tscr_patient_sample.csv")
```

```{r}

```
