---
title: "Tertiary lymphoid structure in MCC"
output: html_notebook
---

```{r}
suppressMessages({
  library(Seurat)
  library(ggplot2)
  library(future)
  library(RColorBrewer)
  library(SingleCellExperiment)
  library(BayesSpace)
  library(dplyr)
})
```

```{r}
options(future.globals.maxSize = 50*1024*1024^2)
set.seed(100)
`%!in%` <- Negate(`%in%`)
source("utils/plot_func.R")
```

## TLS analysis with Me31 from Weng-Onn, using BayesSpace data

```{r}
seurats <- readRDS("~/data/wengonn/seurat_obs_karolinska_all.rds")
me31 <- seurats[[2]]
me31
```

```{r}
SpatialDimPlot(me31)
DimPlot(me31)
```

```{r}

```

Read data from disk

```{r}
me31.bayes <- readRDS("~/data/wengonn/bayesspace_Me31_outs.rds")
# me31.ffpe.bayes <- readRDS("~/data/951572_MCC_Mel_Visium/bayesspace_Me31_MCC_outs.rds")
me31.bayes
```

Flip coordinates (optional, depends on orientation in Seurat)

```{r}
me31.bayes$sce.enhanced$spot.col <- -me31.bayes$sce.enhanced$spot.col
me31.bayes$sce$col <- -me31.bayes$sce$col
```

```{r}
colors <- rbind(RColorBrewer::brewer.pal(8, name = "Set2"), RColorBrewer::brewer.pal(8, name = "Dark2"))
colors <- c(colors) # Linearize
colors <- c(colors, brewer.pal(12, name = "Set3"))
names(colors) <- paste("C",seq(length(colors)),sep="")

clusterPlot(me31.bayes$sce.enhanced) + scale_fill_manual(values = unname(colors))
qPlot(me31.bayes$sce)
logcounts(me31.bayes$sce.enhanced) %>% dim
```

Create Seurat object with logcounts from BayesSpace

```{r}
# so.for.umap <- CreateSeuratObject(counts = logcounts(me31.bayes$sce.enhanced))
# so.for.umap[["RNA"]]$data <- so.for.umap[["RNA"]]$counts
# so.for.umap$label <- me31.bayes$sce.enhanced$spatial.cluster
# so.for.umap
# 
# so.for.umap <- FindVariableFeatures(so.for.umap)
# so.for.umap <- ScaleData(so.for.umap)
# so.for.umap <- RunPCA(so.for.umap)
# so.for.umap <- RunUMAP(so.for.umap, dims=1:10, verbose=F)
# so.for.umap <- RunTSNE(so.for.umap, dims=1:10, verbose=F)
# 
# saveRDS(so.for.umap, "~/data/wengonn/seurat_me31_bayesspace_logcounts.rds")
so.for.umap <- readRDS("~/data/wengonn/seurat_me31_bayesspace_logcounts.rds")

DimPlot(so.for.umap, group.by="label") + xlab("UMAP 1") + ylab("UMAP 2") + ggtitle("") + scale_color_manual(values = unname(colors))
DimPlot(so.for.umap, group.by="label", reduction = "tsne") + xlab("TSNE 1") + ylab("TSNE 2") + ggtitle("") + scale_color_manual(values = unname(colors))
```

Assign Seurat clusters into tumor group or immune group

```{r}
so.for.umap$group <- plyr::mapvalues(so.for.umap$label, from=c(1,2,6,7,5,3,4), 
                                     to=c(rep("Immune",3),rep("Tumor",4),rep(NA,0)))
DimPlot(so.for.umap, group.by="group", cols=brewer.pal(8, "Set1")) + xlab("UMAP 1") + ylab("UMAP 2")

me31.bayes$sce.enhanced$group <- so.for.umap$group
clusterPlot(me31.bayes$sce.enhanced, label = "group") + scale_fill_manual(values = brewer.pal(8, "Set1"))
```

Determine tumor-TLS edge based on CHGA expression

```{r}
so.for.umap.tls <- subset(so.for.umap, group=="Immune")
so.for.umap.tls
chga.thres <- quantile(so.for.umap.tls[["RNA"]]$data["CHGA",], 0.7)
chga.thres
tls.tumor.idents <- plyr::mapvalues(so.for.umap.tls[["RNA"]]$data["CHGA",] > chga.thres, from=c(TRUE, FALSE), to=c("Tumor-TLS", "TLS"))
so.for.umap$group2 <- plyr::mapvalues(colnames(so.for.umap), from=names(tls.tumor.idents), to=tls.tumor.idents)
so.for.umap$group2 <- factor(gsub("subspot_[0-9\\.]+", "Tumor", so.for.umap$group2), levels=c("TLS","Tumor-TLS","Tumor"))
table(so.for.umap$group2)
```

```{r}
me31.bayes$sce.enhanced$group2 <- factor(so.for.umap$group2, levels=c("TLS","Tumor-TLS","Tumor"))
custom.cols <- brewer.pal(8, "Set1")[c(1,3,2)]
clusterPlot(me31.bayes$sce.enhanced, label = "group2") + scale_fill_manual(values = custom.cols)
DimPlot(so.for.umap, group.by="group2", cols=custom.cols) + xlab("UMAP 1") + ylab("UMAP 2") + ggtitle("")
```

Check gene expression in tumor-TLS regions

```{r}
l <- VlnPlot(so.for.umap, c("OAS2","IFI27","MX1","PARP14","PARP9","IRF9"), pt.size=0, group.by="group2", col=custom.cols, combine=FALSE)
patchwork::wrap_plots(lapply(l, function(x) x+xlab("") + ylab("") + theme(legend.position="none", plot.title = element_text(size=15))), ncol=3)

l <- VlnPlot(so.for.umap, c("CCL21","CCR7"), pt.size=0, group.by="group2", col=custom.cols, combine=FALSE)
patchwork::wrap_plots(lapply(l, function(x) x+coord_cartesian(ylim=c(0,1))+xlab("") + ylab("") + theme(legend.position="none", plot.title = element_text(size=15))), ncol=4)

l <- VlnPlot(so.for.umap, c("IGHG1","IGHM","PTGDS","JCHAIN","IRF7","IRF8"), pt.size=0, group.by="group2", col=custom.cols, combine=FALSE)
patchwork::wrap_plots(lapply(l, function(x) x+coord_cartesian(ylim=c(0,1))+xlab("") + ylab("") + theme(legend.position="none", plot.title = element_text(size=15))), ncol=4)
```

Check gene expression in spatial coordinates

```{r}
plot.bayesspace(me31.bayes$sce.enhanced, me31.bayes$sce, c("OAS2","IFI27","MX1","PARP14","PARP9","IRF9"), 3)
plot.bayesspace(me31.bayes$sce.enhanced, me31.bayes$sce, c("IGHG1","IGHM","PTGDS","JCHAIN","IRF7","IRF8"), 3)
```

```{r}
sessionInfo()
```
